The sound of teaching music: Expert pianists' performance modulations for novices

### 2022-09-01 paper/manuscript.Rmd
add captions for tables / correct the wrong plot for vel_dyn_1.png
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

# save plots
ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 300, width = 7, height = 5)
img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"), density = 300)

# annotate * for forte
img_vel_dyn_1 <- image_annotate(img_vel_dyn_1, "*", size = 130, color = "black", location = "+725+400")
```

@@

Musicians and music teachers tend to consider expressive skills as performers’ most important skills [@lindstrom_2003]. In the present study, experts focused on the teaching of expressions that were notated. However, expression also has other facets [@juslin_2003] that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer’s interpretation, listening contexts such as recordings or concerts). The ultimate goal of expressive performance is considered to lie in developing performers’ own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing [@meissner_2021]. The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [e.g., @woody_2000]. It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.

@@

apa_table(
  ioi_all
  , caption = "Experiment 1:"
  , note = "Note"
  , escape = TRUE
)
```

@@

apa_table(
  kot_art_all
  , caption = "Experiment 1:"
  , note = "Note"
  , escape = TRUE
)
```

@@

apa_table(
  vel_dyn_all
  , caption = "Experiment 1:"
  , note = "Note"
  , escape = TRUE
)
```

@@

apa_table(
  vel_diff_dyn_all
  , caption = "Experiment 1:"
  , note = "Note"
  , escape = TRUE
)
```

@@

```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
img_ioi_teaching_2 <- image_append(c(img_ioi_art_teaching_2, img_ioi_dyn_teaching_2))
image_write(img_ioi_teaching_2, path = here("paper/figure", "Fig17.tiff"), format = "tiff")
img_ioi_teaching_2
```

```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
img_art_teaching_2 <- image_append(c(img_art_leg_teaching_2, img_art_sta_teaching_2))
image_write(img_art_teaching_2, path = here("paper/figure", "Fig18.tiff"), format = "tiff")
img_art_teaching_2
```

@@

apa_table(
  ioi_all
  , caption = "Experiment 2:"
  , note = "Note"
  , escape = TRUE
  , digits = 3
)
```

@@

apa_table(
  kor_art_all
  , caption = "Experiment 2:"
  , note = "Note"
  , escape = TRUE
  , digits = 3
)
```

@@

apa_table(
  vel_dyn_all
  , caption = "Experiment 2:"
  , note = "Note"
  , escape = TRUE
)
```

@@

apa_table(
  vel_diff_dyn_all
  , caption = "Experiment 2:"
  , note = "Note"
  , escape = TRUE
)
```

>>>>

# save plots
ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 300, width = 7, height = 5)
img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"), density = 300)
```

@@

Musicians and music teachers tend to consider expressive skills as performers’ most important skills [@lindstrom_2003]. In the present study, experts focused on the teaching of expressions that were notated. However, expression also has other facets [@juslin_2003] that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer’s interpretation, listening contexts such as recordings or concerts). The ultimate goal of expressive performance is considered to lie in developing performers’ own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing [@meissner_2021]. The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [@woody_2000; @colprit_2000]. Given that teachers demonstrate for students [@colprit_2000] and students often learn by listening to the performances of either their teachers or recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.

@@

apa_table(
  ioi_all[, -c(8, 9)]
  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms)."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```

@@

apa_table(
  kot_art_all
  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms)."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```

@@

apa_table(
  vel_dyn_all
  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127)."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```

@@

apa_table(
  vel_diff_dyn_all
  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```

@@

```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs when playing the piece with either articulation (left) or dynamics (right). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
img_ioi_teaching_2 <- image_append(c(img_ioi_art_teaching_2, img_ioi_dyn_teaching_2))
image_write(img_ioi_teaching_2, path = here("paper/figure", "Fig17.tiff"), format = "tiff")
img_ioi_teaching_2
```

```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
img_art_teaching_2 <- image_append(c(img_art_leg_teaching_2, img_art_sta_teaching_2))
image_write(img_art_teaching_2, path = here("paper/figure", "Fig18.tiff"), format = "tiff")
img_art_teaching_2
```

@@

apa_table(
  ioi_all[, -c(8, 9)]
  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
  , digits = 3
)
```

@@

apa_table(
  kor_art_all
  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KOR."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
  , digits = 3
)
```

@@

apa_table(
  vel_dyn_all
  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127)."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```

@@

apa_table(
  vel_diff_dyn_all
  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points."
  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano."
  , escape = TRUE
)
```
